\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#662}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, distances

\medskip
\noindent\textbf{Note:} ambiguous statement

\medskip

\noindent\textbf{Problem Statement:}

Consider the triangular lattice with minimal distance between two points $1$. Denote by $f(t)$ the number of distances from any points $\leq t$. For example $f(1)=6$, $f(\sqrt{3})=12$, and $f(3)=18$. Let $x_1,\ldots,x_n\in \mathbb{R}^2$ be such that $d(x_i,x_j)\geq 1$ for all $i\neq j$. Is it true that, provided $n$ is sufficiently large depending on $t$, the number of distances $d(x_i,x_j)\leq t$ is less than or equal to $f(t)$ with equality perhaps only for the triangular lattice? In particular, is it true that the number of distances $\leq \sqrt{3}-\epsilon$ is less than $1$?




A problem of Erd\H{o}s, Lov\'{a}sz, and Vesztergombi. This is essentially verbatim the problem description in \cite{Er97e}, but this does not make sense as written; there must be at least one typo. Suggestions about what this problem intends are welcome. Erd\H{o}s also goes on to write 'Perhaps the following stronger conjecture holds: Let $t_1&#60;t_2&#60;\cdots$ be the set of distances occurring in the triangular lattice. $t_1=1$ $t_2=\sqrt{3}$ $t_3=3$ $t_4=5$ etc. Is it true that there is an $\epsilon_n$ so that for every set $y_1,\ldots,$ with $d(y_i,y_j)\geq 1$ the number of distances $d(y_i,y_j)&#60;t_n$ is less than $f(t_n)$?' Again, this is nonsense interpreted literally; I am not sure what Erd\H{o}s intended.




References

[Er97e] Erd\H{o}s, Paul, Some of my favourite unsolved problems . Math. Japon. (1997), 527-537.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/662}}
\end{document}
