\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1084}
\date{Last updated: 2025-10-17}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, distances

\medskip
\noindent\textbf{Note:} contact number problem

\medskip

\noindent\textbf{Problem Statement:}

Let $f_d(n)$ be minimal such that in any collection of $n$ points in $\mathbb{R}^d$, all of distance at least $1$ apart, there are at most $f_d(n)$ many pairs of points which are distance $1$ apart. Estimate $f_d(n)$.




This is sometimes known as the contact number problem. It is easy to see that $f_1(n)=n-1$ and $f_2(n)&lt;3n$ (since there can be at most $6$ points of distance $1$ from any point). Erd\H{o}s \cite{Er46b} showed\[f_2(n)&lt;3n-cn^{1/2}\]for some constant $c&gt;0$, which the triangular lattice shows is the best possible up to the value of $c$. In \cite{Er75f} he speculated that the triangular lattice is exactly the best possible, and in particular\[f_2(3n^2+3n+1)=9n^2+3n.\]Harborth \cite{Ha74b} proved this, and more generally\[f_2(n)=\lfloor 3n-\sqrt{12n-3}\rfloor\]for all $n\geq 2$. In \cite{Er75f} he claims the existence of $c_1,c_2&gt;0$ such that\[6n-c_1n^{2/3}&lt; f_3(n) &lt; 6n-c_2n^{2/3}.\]An upper bound of\[f_3(n) &lt; 6n-0.926n^{2/3}\]for all $n\geq 2$ was proved by Bezdek and Reid \cite{BeRe13}. In general, it is known that\[(d-o(1))n \leq f_d(n) \leq 2^{O(d)}n,\]the lower bound coming from points arranged in an integer grid and the upper bound from the fact that $2^{O(d)}$ many non-intersecting congruent balls can touch a fixed ball (the kissing number problem). A recent survey on contact numbers for sphere packings is by Bezdek and Khan \cite{BeKh18}. See [223] for the analogous problem with maximal distance $1$.




References

[BeKh18] Bezdek, K\'{a}roly and Khan, Muhammad A., Contact numbers for sphere packings . (2018), 25--47.

[BeRe13] Bezdek, K\'{a}roly and Reid, Samuel, Contact graphs of unit sphere packings revisited . J. Geom. (2013), 57--83.

[Er46b] Erd\H{o}s, P., On sets of distances of {$n$} points . Amer. Math. Monthly (1946), 248--250.

[Er75f] Erd\H{o}s, Paul, On some problems of elementary and combinatorial geometry . Ann. Mat. Pura Appl. (4) (1975), 99-108.

[Ha74b] Harborth, Heiko, L\"{o}sung zu Problem 664A . Elem. Math. (1974), 14-15.

\noindent\textbf{Related OEIS sequences:} A045945, possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1084}}
\end{document}
