\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1045}
\date{Last updated: 2026-03-14}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} analysis

\medskip

\noindent\textbf{Problem Statement:}

Let $z_1,\ldots,z_n\in \mathbb{C}$ with $\lvert z_i-z_j\rvert\leq 2$ for all $i,j$, and\[\Delta(z_1,\ldots,z_n)=\prod_{i\neq j}\lvert z_i-z_j\rvert.\]What is the maximum possible value of $\Delta$? Is it maximised by taking the $z_i$ to be the vertices of a regular polygon?




A problem of Erd\H{o}s, Herzog, and Piranian \cite{EHP58}, who proved that, for any monic polynomial $f$, if $\{ z: \lvert f(z)\rvert &lt;1\}$ is connected and $f$ has roots $z_1,\ldots,z_n$ then $\prod_{i\neq j}\lvert z_i-z_j\rvert &lt;n^n$. The value of $\Delta$ when the $z_i$ are the vertices of a regular polygon is $n^n$ when $n$ is even and\[\cos(\pi/2n)^{-n(n-1)}n^n \sim e^{\pi^2/8}n^n\]when $n$ is odd. Pommerenke \cite{Po61} proved that $\Delta \leq 2^{O(n)}n^n$ for all $z_i$ with $\lvert z_i-z_j\rvert \leq 2$. Hu and Tang (see the comments) found examples when $n=4$ and $n=6$ that show that vertices of a regular polygon do not maximise $\Delta$. Cambie (also in the comments) showed that, in general, the vertices of a regular polygon are not a maximiser for all even $n\geq 4$. There is a lot of discussion of this problem in the comments (see also the papers of Sothanaphan \cite{So25} and Cambie, Decadt, Dong, Hu, and Tang \cite{CDDHT26}). It is now known that, for even $n$,\[\liminf \frac{\max \Delta}{n^n}\geq C\]for some $C&gt;0$. The best value of $C$ available is $C\approx 1.268$, proved in \cite{CDDHT26}. If we restrict to $n$ divisible by $6$ this can be improved to $C\approx 1.304$. It remains possible that the regular polygon is a maximiser for odd $n$; this is explicitly conjectured in \cite{CDDHT26}. It is not yet known whether, for odd $n$,\[\lim \frac{\max \Delta}{n^n}=e^{\pi^2/8}\approx 3.433.\]The paper \cite{CDDHT26} also contains a number of structural results about the form of an extremal set of points.




References

[CDDHT26] S. Cambie, A. Decadt, Y. Dong, T. Hu, and Q. Tang, On the maximum product of distances of diameter 2 point sets . arXiv:2603.07088 (2026).

[EHP58] Erd\H{o}s, P. and Herzog, F. and Piranian, G., Metric properties of polynomials . J. Analyse Math. (1958), 125-148.

[Po61] Pommerenke, Ch., On metric properties of complex polynomials . Michigan Math. J. (1961), 97-115.

[So25] N. Sothanaphan, An improved lower bound to Erdos' problem concerning products of distances for fixed diameter . arXiv:2512.14251 (2025).


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1045}}
\end{document}
