\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#827}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry

\medskip

\noindent\textbf{Problem Statement:}

Let $n_k$ be minimal such that if $n_k$ points in $\mathbb{R}^2$ are in general position then there exists a subset of $k$ points such that all $\binom{k}{3}$ triples determine circles of different radii. Determine $n_k$.




In \cite{Er75h} Erd\H{o}s asks whether $n_k$ exists. In \cite{Er78c} he gave a simple argument which proves that it does, and in fact\[n_k \leq k+2\binom{k-1}{2}\binom{k-1}{3},\]but this argument is incorrect, as explained by Martinez and Rold\'{a}n-Pensado \cite{MaRo15}. Martinez and Rold\'{a}n-Pensado give a corrected argument that proves $n_k\ll k^9$. A simple probabilistic argument, given in the comments by FlaredRain, improves this to $n_k\ll k^5$.




References

[Er75h] Erd\H{o}s, P., Some problems on elementary geometry . Austral. Math. Soc. Gaz. (1975), 2-3.

[Er78c] Erd\H{o}s, P., Some more problems on elementary geometry . Austral. Math. Soc. Gaz. (1978), 52-54.

[MaRo15] Mart\'Inez, L. and Rold\'{a}n-Pensado, E., Points defining triangles with distinct circumradii . Acta Math. Hungar. (2015), 136--141.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/827}}
\end{document}
