\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#483}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, additive combinatorics, ramsey theory

\medskip

\noindent\textbf{Problem Statement:}

Let $f(k)$ be the minimal $N$ such that if $\{1,\ldots,N\}$ is $k$-coloured then there is a monochromatic solution to $a+b=c$. Estimate $f(k)$. In particular, is it true that $f(k) &lt; c^k$ for some constant $c&gt;0$?




The values of $f(k)$ are known as Schur numbers. The best-known bounds for large $k$ are\[(380)^{k/5}-O(1)\leq f(k) \leq (e-\tfrac{1}{6}) k!.\]Note that $380^{1/5}\approx 3.2806$. The lower bound is due to Ageron, Casteras, Pellerin, Portella, Rimmel, and Tomasik \cite{ACPPRT21} (improving previous bounds of Exoo \cite{Ex94} and Fredricksen and Sweet \cite{FrSw00}). The upper bound is due to Xu, Xie, and Chen \cite{XXC02} (improving previous bounds of Wan \cite{Wa97} and Whitehead \cite{Wh73}). See Eliahou \cite{El19} for more on the upper bound. The known values of $f$ are $f(1)=2$, $f(2)=5$, $f(3)=14$, $f(4)=45$, and $f(5)=161$ (see A030126 ). (The equality $f(5)=161$ was established by Heule \cite{He17}). See also [183] (in particular a folklore observation gives $f(k)\leq R(3;k)-1$).




References

[ACPPRT21] R. Ageron, P. Casteras, T. Pellerin, Y. Portella, A. Rimmel, and J. Tomasik, New lower bounds for Schur and weak Schur numbers . arXiv:2112.03175 (2021).

[El19] No reference found.


[Ex94] Exoo, G., A lower bound for Schur numbers and multicolor Ramsey numbers . Electronic J. of Combinatorics (1994).

[FrSw00] Fredricksen, Harold and Sweet, Melvin M., Symmetric sum-free partitions and lower bounds for Schur numbers . Electron. J. Combin. (2000), Research Paper 32, 9.

[He17] M. Heuele, Schur Number Five . arXiv:1711.08076 (2017).

[Wa97] Wan, Honghui, Upper bounds for {R}amsey numbers {$R(3,3,\cdots,3)$} and Schur numbers . J. Graph Theory (1997), 119--122.

[Wh73] Whitehead, Jr., Earl Glen, The {R}amsey number {$N(3,\,3,\,3,\,3;\,2)$} . Discrete Math. (1973), 389--396.

[XXC02] Xu, Xiao Dong and Xie, Zheng and Chen, Zhi, Upper bounds for {R}amsey numbers {$R_n(3)$} and Schur numbers . Math. Econ. (2002), 81--84.

\noindent\textbf{Related OEIS sequences:} A030126


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/483}}
\end{document}
