\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#583}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} FALSIFIABLE \quad \textbf{No prize}

\noindent\textbf{Tags:} graph theory

\medskip

\noindent\textbf{Problem Statement:}

Every connected graph on $n$ vertices can be partitioned into at most $\lceil n/2\rceil$ edge-disjoint paths.




A problem of Erd\H{o}s and Gallai. If we drop the edge-disjoint condition then this conjecture was proved by Fan \cite{Fa02}. There are many partial results towards this conjecture. We list some highlights below; $G$ is an arbitrary connected graph on $n$ vertices. {UL} {LI} Lov\'{a}sz \cite{Lo68} proved that $G$ can be partitioned into at most $\lfloor n/2\rfloor$ edge-disjoint paths and cycles, and hence into at most $n-1$ paths. This also implies the original conjecture if $G$ has at most one vertex of even degree. {/LI} {LI} Chung \cite{Ch78} proved that $G$ can be partitioned into at most $\lceil n/2\rceil$ edge-disjoint trees.{/LI} {LI} Pyber \cite{Py96} proved that $G$ can be covered by at most $n/2+O(n^{3/4})$ paths (that may not be edge-disjoint).{/LI} {LI} Pyber \cite{Py96} also proved the original conjecture if the subgraph induced by vertices of even degree is a forest.{/LI} {LI} Bonamy and Perrett \cite{BoPe19} proved the conjecture if $G$ has maximum degree $\leq 5$.{/LI} {LI} Blanch\'{e}, Bonamy, and Bonichon \cite{BBB21} proved this conjecture if $G$ is planar.{/LI} {LI} Anto and Basavaraju \cite{AnBa23} proved the conjecture if $G$ is $2$-degenerate (every subgraph has a vertex of degree $\leq 2$).{/LI} {LI} Chu, Fan, and Zhou \cite{CFZ26} proved the conjecture if the subgraph induced by vertices of even degree is $K_m$ for some $m\leq 15$ and $n$ is odd.{/LI} {LI} Dean and Kouider \cite{DeKo00} prove that $\lceil \frac{2}{3}n\rceil$ edge-disjoint paths always suffice, even for disconnected graphs (for which this bound is optimal). The same bound was independently obtained by Yan in their 1998 PhD thesis .{/LI} {/UL} Haj\'{o}s \cite{Lo68} has conjectured that if $G$ has all degrees even then $G$ can be partitioned into at most $\lfloor n/2\rfloor$ edge-disjoint cycles. See also [184] for an analogous problem decomposing into edges and cycles and [1017] for decomposing into complete graphs. See also the entry in the graphs problem collection .




References

[AnBa23] Anto, Nevil and Basavaraju, Manu, Gallai's path decomposition for 2-degenerate graphs . Discrete Math. Theor. Comput. Sci. (2023), Paper No. 16, 11.

[BBB21] A. Blanch\'{e}, M. Bonamy, and N. Bonichon, Gallai's path decomposition in planar graphs . arXiv:2110.08870 (2021).

[BoPe19] Bonamy, Marthe and Perrett, Thomas J., Gallai's path decomposition conjecture for graphs of small maximum degree . Discrete Math. (2019), 1293--1299.

[CFZ26] Chu, Yanan and Fan, Genghua and Zhou, Chuixiang, Gallai's conjecture and the path number of odd semi-cliques . Discrete Math. (2026), Paper No. 114725, 6.

[Ch78] Chung, F. R. K., On partitions of graphs into trees . Discrete Math. (1978), 23-30.

[DeKo00] Dean, Nathaniel and Kouider, Mekkia, Gallai's conjecture for disconnected graphs . Discrete Math. (2000), 43--54.

[Fa02] Fan, Genghua, Subgraph coverings and edge switchings . J. Combin. Theory Ser. B (2002), 54-83.

[Lo68] Lov\'{a}sz, L., On covering of graphs . Theory of Graphs (Proc. Colloq., Tihany, 1966) (1968), 231-236.

[Py96] Pyber, L., Covering the edges of a connected graph by paths . J. Combin. Theory Ser. B (1996), 152-159.

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/583}}
\end{document}
