\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#743}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} FALSIFIABLE \quad \textbf{No prize}

\noindent\textbf{Tags:} graph theory

\medskip
\noindent\textbf{Note:} tree packing conjecture

\medskip

\noindent\textbf{Problem Statement:}

Let $T_2,\ldots,T_n$ be a collection of trees such that $T_k$ has $k$ vertices. Can we always write $K_n$ as the edge disjoint union of the $T_k$?




A conjecture of Gy\'{a}rf\'{a}s, known as the tree packing conjecture. Gy\'{a}rf\'{a}s and Lehel \cite{GyLe78} proved that this holds if all but at most $2$ of the trees are stars, or if all the trees are stars or paths. Fishburn \cite{Fi83} proved this for $n\leq 9$. Bollob\'{a}s \cite{Bo83} proved that the smallest $\lfloor n/\sqrt{2}\rfloor$ many trees can always be packed greedily into $K_n$. Joos, Kim, K\"{u}hn, and Osthus \cite{JKKO19} proved that this conjecture holds when the trees have bounded maximum degree. Allen, B\"{o}ttcher, Clemens, Hladky, Piguet, and Taraz \cite{ABCHPT21} proved that this conjecture holds when all the trees have maximum degree $\leq c\frac{n}{\log n}$ for some constant $c&#62;0$. Janzer and Montgomery \cite{JaMo24} have proved that there exists some $c&#62;0$ such that the largest $cn$ trees can be packed into $K_n$.




References

[ABCHPT21] Allen, Peter and B\"ottcher, Julia and Clemens, Dennis and Hladky, J. and D. Piguet and Taraz, Anusch, The tree packing conjecture for trees of almost linear maximum degree . arXiv:2106.11720 (2021).

[Bo83] Bollob\'{a}s, B\'{e}la, Some remarks on packing trees . Discrete Math. (1983), 203-204.

[Fi83] Fishburn, P. C., Balanced integer arrays: a matrix packing theorem . J. Combin. Theory Ser. A (1983), 98-101.

[GyLe78] Gy\'{a}rf\'{a}s, A. and Lehel, J., Packing trees of different order into {$K\sb{n}$} . (1978), 463-469.

[JKKO19] Joos, Felix and Kim, Jaehoon and K\"{u}hn, Daniela and Osthus, Deryk, Optimal packings of bounded degree trees . J. Eur. Math. Soc. (JEMS) (2019), 3573-3647.

[JaMo24] Janzer, B. and R. Montgomery, Packing the largest trees in the tree packing conjecture . arXiv:2403.10515 (2024).

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/743}}
\end{document}
