\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#128}
\date{Last updated: 2025-10-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} FALSIFIABLE \quad \textbf{Prize: \$250}

\noindent\textbf{Tags:} graph theory

\medskip

\noindent\textbf{Problem Statement:}

Let $G$ be a graph with $n$ vertices such that every induced subgraph on $\geq \lfloor n/2\rfloor$ vertices has more than $n^2/50$ edges. Must $G$ contain a triangle?




A problem of Erd\H{o}s and Rousseau. The constant $50$ would be best possible as witnessed by a blow-up of $C_5$ or the Petersen graph. Erd\H{o}s, Faudree, Rousseau, and Schelp \cite{EFRS94} proved that this is true with $50$ replaced by $16$. More generally, they prove that, for any $0&lt;\alpha&lt;1$, if every set of $\geq \alpha n$ vertices contains $&gt;\alpha^3n^2/2$ edges then $G$ contains a triangle. Krivelevich \cite{Kr95} has proved this with $n/2$ replaced by $3n/5$ (and $50$ replaced by $25$). Keevash and Sudakov \cite{KeSu06} have proved this under the additional assumption that either $G$ has at most $n^2/12$ edges, or that $G$ has at least $n^2/5$ edges. Norin and Yepremyan \cite{NoYe15} proved that this is true if $G$ has at least $(1/5-c)n^2$ edges, for some constant $c&gt;0$. Razborov \cite{Ra22} proved this is true if $\frac{1}{50}$ is replaced by $\frac{27}{1024}$. See also the entry in the graphs problem collection .




References

[EFRS94] Erd\H{o}s, P. and Faudree, R. J. and Rousseau, C. C. and Schelp, R. H., A local density condition for triangles . Discrete Math. (1994), 153--161.

[KeSu06] Keevash, Peter and Sudakov, Benny, Sparse halves in triangle-free graphs . J. Combin. Theory Ser. B (2006), 614-620.

[Kr95] Krivelevich, Michael, On the edge distribution in triangle-free graphs . J. Combin. Theory Ser. B (1995), 245-260.

[NoYe15] Norin, Sergey and Yepremyan, Liana, Sparse halves in dense triangle-free graphs . J. Combin. Theory Ser. B (2015), 1--25.

[Ra22] Razborov, A. A., More about sparse halves in triangle-free graphs . Mat. Sb. (2022), 119--140.

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/128}}
\end{document}
