\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#107}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} FALSIFIABLE \quad \textbf{Prize: \$500}

\noindent\textbf{Tags:} geometry, convex

\medskip
\noindent\textbf{Note:} 'Happy Ending' problem

\medskip

\noindent\textbf{Problem Statement:}

Let $f(n)$ be minimal such that any $f(n)$ points in $\mathbb{R}^2$, no three on a line, contain $n$ points which form the vertices of a convex $n$-gon. Prove that $f(n)=2^{n-2}+1$.




The Erd\H{o}s-Klein-Szekeres 'Happy Ending' problem. The problem originated in 1931 when Klein observed that $f(4)=5$. Tur\'{a}n and Makai showed $f(5)=9$. Erd\H{o}s and Szekeres proved the bounds\[2^{n-2}+1\leq f(n)\leq \binom{2n-4}{n-2}+1.\](\cite{ErSz60} and \cite{ErSz35} respectively). There were several improvements of the upper bound, but all of the form $4^{(1+o(1))n}$, until Suk \cite{Su17} proved\[f(n) \leq 2^{(1+o(1))n}.\]The current best bound is due to Holmsen, Mojarrad, Pach, and Tardos \cite{HMPT20}, who prove\[f(n) \leq 2^{n+O(\sqrt{n\log n})}.\]In \cite{Er97e} Erd\H{o}s clarifies that the \$500 is for a proof, and only offers \$100 for a disproof. Graham \cite{Gr04} offers \$1000 for a proof. This problem is #1 in Ramsey Theory in the graphs problem collection. See also [216] , [651] , and [838] .




References

[Er97e] Erd\H{o}s, Paul, Some of my favourite unsolved problems . Math. Japon. (1997), 527-537.

[ErSz35] Erd\H{o}s, P. and Szekeres, G., A combinatorial problem in geometry . Compos. Math. (1935), 463-470.

[ErSz60] Erd\H{o}s, P. and Szekeres, G., On some extremum problems in elementary geometry . Ann. Univ. Sci. Budapest. E\"{o}tv\"{o}s Sect. Math. (1960/61), 53-62.

[Gr04] Green, Ben, The Cameron-Erd\H{o}s conjecture . Bull. London Math. Soc. (2004), 769-778.

[HMPT20] Holmsen, Andreas F. and Mojarrad, Hossein Nassajian and Pach, J\'{a}nos and Tardos, G\'{a}bor, Two extensions of the Erd\H{o}s-Szekeres problem . J. Eur. Math. Soc. (JEMS) (2020), 3981-3995.

[Su17] Suk, Andrew, On the Erd\H{o}s-Szekeres convex polygon problem . J. Amer. Math. Soc. (2017), 1047-1053.

\noindent\textbf{Related OEIS sequences:} A000051

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/107}}
\end{document}
