\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#560}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} graph theory, ramsey theory

\medskip

\noindent\textbf{Problem Statement:}

Let $\hat{R}(G)$ denote the size Ramsey number, the minimal number of edges $m$ such that there is a graph $H$ with $m$ edges such that in any $2$-colouring of the edges of $H$ there is a monochromatic copy of $G$. Determine\[\hat{R}(K_{n,n}),\]where $K_{n,n}$ is the complete bipartite graph with $n$ vertices in each component.




We know that\[\frac{1}{60}n^22^n&lt;\hat{R}(K_{n,n})&lt; \frac{3}{2}n^32^n.\]The lower bound (which holds for $n\geq 6$) was proved by Erd\H{o}s and Rousseau \cite{ErRo93}. The upper bound was proved by Erd\H{o}s, Faudree, Rousseau, and Schelp \cite{EFRS78b} and Ne\v{s}et\v{r}il and R\"{o}dl \cite{NeRo78}. Conlon, Fox, and Wigderson \cite{CFW23} have proved that, for any $s\leq t$,\[\hat{R}(K_{s,t})\gg s^{2-\frac{s}{t}}t2^s,\]and prove that when $t\gg s\log s$ we have $\hat{R}(K_{s,t})\asymp s^2t2^s$. They conjecture that this should hold for all $s\leq t$, and so in particular we should have $\hat{R}(K_{n,n})\asymp n^32^n$. This problem is #29 in Ramsey Theory in the graphs problem collection.




References

[CFW23] Conlon, David and Fox, Jacob and Wigderson, Yuval, Three early problems on size Ramsey numbers . Combinatorica (2023), 743-768.

[EFRS78b] Erd\H{o}s, P. and Faudree, R. J. and Rousseau, C. C. and Schelp, R. H., The size Ramsey number . Period. Math. Hungar. (1978), 145-161.

[ErRo93] Erd\H{o}s, P. and Rousseau, C. C., The size Ramsey number of a complete bipartite graph . Discrete Math. (1993), 259-262.

[NeRo78] Ne\vSet\v{r}il, J. and R\"{o}dl, V., The structure of critical Ramsey graphs . Acta Math. Acad. Sci. Hungar. (1978), 295-300.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/560}}
\end{document}
