\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#371}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory

\medskip

\noindent\textbf{Problem Statement:}

Let $P(n)$ denote the largest prime factor of $n$. Show that the set of $n$ with $P(n)&#60;P(n+1)$ has density $1/2$.




Conjectured by Erd\H{o}s and Pomerance \cite{ErPo78}, who proved that this set and its complement both have positive upper density. The best unconditional lower bound available is due to L\"{u} and Wang \cite{LuWa25}, who prove that\[\#\{ n&lt;x :P(n)&lt;P(n+1)\} &gt; (0.2017-o(1))x,\]and the same lower bound for the complement. In \cite{Er79e} Erd\H{o}s also asks whether, for every $\alpha$, the density of the set of $n$ where\[P(n+1)&gt;P(n)n^\alpha\]exists. Ter\"{a}v\"{a}inen \cite{Te18} has proved that the logarithmic density of the set of $n$ for which $P(n)&lt;P(n+1)$ is $1/2$. Tao and Ter\"{a}v\"{a}inen \cite{TaTe19} have proved that the asymptotic density is $1/2$ at 'almost all scales'. More generally, for any $0\leq \alpha \leq1$, Ter\"{a}v\"{a}inen \cite{Te18} proved that the logarithmic density of the set of $n$ for which $P(n+1)&gt;P(n)n^\alpha$ exists and is equal to\[\int_{[0,1]^2}1_{y\geq x+\alpha}u(x)u(y)\mathrm{d}x\mathrm{d}y\]where $u(x)=x^{-1}\rho(x^{-1}-1)$ and $\rho$ is the Dickman function . Wang \cite{Wa21} has proved the same value holds for the asymptotic density (and in particular provided an affirmative answer to the original question) conditional on the Elliott-Halberstam conjecture for friable integers. The sequence of such $n$ is A070089 in the OEIS. See also [372] and [928] .




References

[Er79e] Erd\H{o}s, Paul, Some unconventional problems in number theory . Ast\'{e}risque (1979), 73-82.

[ErPo78] Erd\H{o}s, Paul and Pomerance, Carl, On the largest prime factors of {$n$} and {$n+1$} . Aequationes Math. (1978), 311-321.

[LuWa25] L\"u, Xiaodong and Wang, Zhiwei, On the largest prime factors of consecutive integers . Monatsh. Math. (2025), 403--418.

[TaTe19] Tao, Terence and Ter\"{a}v\"{a}inen, Joni, The structure of correlations of multiplicative functions at almost all scales, with applications to the {C}howla and {E}lliott conjectures . Algebra Number Theory (2019), 2103--2150.

[Te18] Ter\"{a}v\"{a}inen, Joni, On binary correlations of multiplicative functions . Forum Math. Sigma (2018), Paper No. e10, 41.

[Wa21] Wang, Zhiwei, Three conjectures on {$P^+(n)$} and {$P^+(n+1)$} hold under the {E}lliott-{H}alberstam conjecture for friable integers . J. Number Theory (2021), 1--11.

\noindent\textbf{Related OEIS sequences:} A070089


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/371}}
\end{document}
