\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#503}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, distances

\medskip

\noindent\textbf{Problem Statement:}

What is the size of the largest $A\subseteq \mathbb{R}^d$ such that every three points from $A$ determine an isosceles triangle? That is, for any three points $x,y,z$ from $A$, at least two of the distances $\lvert x-y\rvert,\lvert y-z\rvert,\lvert x-z\rvert$ are equal.




When $d=2$ the answer is $6$ (due to Kelly \cite{ErKe47} - an alternative proof is given by Kov\'{a}cs \cite{Ko24c}). When $d=3$ the answer is $8$ (due to Croft \cite{Cr62}). The best upper bound known in general is due to Blokhuis \cite{Bl84} who showed that\[\lvert A\rvert \leq \binom{d+2}{2}.\]Alweiss has observed a lower bound of $\binom{d+1}{2}$ follows from considering the subset of $\mathbb{R}^{d+1}$ formed of all vectors $e_i+e_j$ where $e_i,e_j$ are distinct coordinate vectors. This set can be viewed as a subset of some $\mathbb{R}^d$, and is easily checked to have the required property. Weisenberg observed in the comments that an additional point can be added to Alweiss' construction, giving a lower bound of $\binom{d+1}{2}+1$. The fact that the truth for $d=3$ is $8$ suggests that neither of these bounds is the truth. See also [1088] for a generalisation.




References

[Bl84] Blokhuis, A., Few-distance sets . (1984), iv+70.

[Cr62] Croft, H. T., $9$-point and $7$-point configurations in $3$-space . Proc. London Math. Soc. (3) (1962), 400-424.

[ErKe47] Erd\H{o}s, Paul and Kelly, L. M., Elementary Problems and Solutions: Solutions: E735 . Amer. Math. Monthly (1947), 227-229.

[Ko24c] Z. Kov\'{a}cs, A note on Erd\H{o}s's mysterious remark . arXiv:2412.05190 (2024).

\noindent\textbf{Related OEIS sequences:} A175769


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/503}}
\end{document}
