\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1083}
\date{Last updated: 2025-10-17}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, distances

\medskip

\noindent\textbf{Problem Statement:}

Let $d\geq 3$, and let $f_d(n)$ be the minimal $m$ such that every set of $n$ points in $\mathbb{R}^d$ determines at least $m$ distinct distances. Estimate $f_d(n)$ - in particular, is it true that\[f_d(n)=n^{\frac{2}{d}-o(1)}?\]




A generalisation of the distinct distance problem [89] to higher dimensions. Erd\H{o}s \cite{Er46b} proved\[n^{1/d}\ll_d f_d(n)\ll_d n^{2/d},\]the upper bound construction being given by a set of lattice points. {UL} {LI} Clarkson, Edelsbrunner, Gubias, Sharir, and Welzl \cite{CEGSW90} proved $f_3(n)\gg n^{1/2}$.{/LI} {LI}Aronov, Pach, Sharir, and Tardos \cite{APST04} proved $f_d(n)\gg n^{\frac{1}{d-90/77}-o(1)}$ for any $d\geq 3$ (for example, $f_3(n)\gg n^{0.546}$).{/LI} {LI}Solymosi and Vu \cite{SoVu08} proved $f_3(n) \gg n^{3/5}$ and\[ f_d(n)\gg_d n^{\frac{2}{d}-\frac{c}{d^2}}\]for all $d\geq 4$ for some constant $c&gt;0$. (The result in their paper for $d=3$ is slightly weaker than stated here, but uses as a black box the bound for distinct distances in $2$ dimensions; we have recorded the consequence of combining their method with the work of Guth and Katz on [89] .){/LI} {/UL} The function $f_d(n)$ is essentially the inverse of the function $g_d(n)$ considered in [1089] - with our definitions, $g_d(n)&gt;m$ if and only if $f_d(m)&lt;n$. The emphasis in this problem is, however, on the behaviour as $d$ is fixed and $n\to \infty$.




References

[APST04] Aronov, Boris and Pach, J\'{a}nos and Sharir, Micha and Tardos, G\'{a}bor, Distinct distances in three and higher dimensions . Combin. Probab. Comput. (2004), 283--293.

[CEGSW90] Clarkson, Kenneth L. and Edelsbrunner, Herbert and Guibas, Leonidas J. and Sharir, Micha and Welzl, Emo, Combinatorial complexity bounds for arrangements of curves and spheres . Discrete Comput. Geom. (1990), 99--160.

[Er46b] Erd\H{o}s, P., On sets of distances of {$n$} points . Amer. Math. Monthly (1946), 248--250.

[SoVu08] Solymosi, J\'ozsef and Vu, Van H., Near optimal bounds for the {E}rd\H{o}s distinct distances problem in high dimensions . Combinatorica (2008), 113--125.

\noindent\textbf{Related OEIS sequences:} A186704, possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1083}}
\end{document}
