\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#811}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} graph theory, ramsey theory

\medskip

\noindent\textbf{Problem Statement:}

Suppose $n\equiv 1\pmod{m}$. We say that an edge-colouring of $K_n$ using $m$ colours is balanced if every vertex sees exactly $\lfloor n/m\rfloor$ many edges of each colours. For which graphs $G$ is it true that, if $m=e(G)$, for all large $n\equiv 1\pmod{m}$, every balanced edge-colouring of $K_n$ with $m$ colours contains a rainbow copy of $G$? (That is, a subgraph isomorphic to $G$ where each edge receives a different colour.)




In \cite{Er91} Erd\H{o}s credits this problem to himself, Pyber, and Tuza. This problem was explored in a paper of Erd\H{o}s and Tuza \cite{ErTu93}. In \cite{Er96} Erd\H{o}s seems to suggest that this might be true for every graph $G$, and specifically asks specific challenge posed in \cite{Er91} and \cite{Er96} is whether, in any balanced edge-colouring of $K_{6n+1}$ by $6$ colours there must exist a rainbow $C_6$ and $K_4$. In general, one can ask for a quantitative version, defining $d_G(n)$ to be minimal (if it exists) such that if $n$ is sufficiently large and the edges of $K_n$ are coloured with $e(G)$ many colours such that the minimum degree of each colour class is $\geq d_G(n)$ then there is a rainbow copy of $G$. Erd\H{o}s and Tuza \cite{ErTu93} proved that\[\lfloor n/6\rfloor \leq d_{C_4}(n) \leq \left(\frac{1}{4}-c\right)n\]for some constant $c&#62;0$. Axenovich and Clemen \cite{AxCl24} have proved that there exist infinitely many graphs without this property. In particular, they show that for any odd $\ell \geq 3$ and $m=\lfloor \sqrt{\ell}+3.5\rfloor$ there exist arbitrarily large $n$ such that $K_n$ has a balanced edge-colouring using $\ell$ colours which contains no rainbow $K_m$. They conjecture that $K_m$ lacks this property for all $m\geq 4$. Clemen and Wagner \cite{ClWa23} proved that $K_4$ does lack this property.




References

[AxCl24] Axenovich, Maria and Clemen, Felix C., Rainbow subgraphs in edge-colored complete graphs: answering two questions by {E}rd\H{o}s and Tuza . J. Graph Theory (2024), 57--66.

[ClWa23] Clemen, Felix Christian and Wagner, Adam Zsolt, Balanced edge-colorings avoiding rainbow cliques of size four . Electron. J. Combin. (2023), Paper No. 3.17, 3.

[Er91] Erd\H{o}s, P., Problems and results in combinatorial analysis and combinatorial number theory . Graph theory, combinatorics, and applications, Vol. 1 (Kalamazoo, MI, 1988) (1991), 397-406.

[Er96] Erd\H{o}s, Paul, Some of my favourite problems on cycles and colourings . Tatra Mt. Math. Publ. (1996), 7-9.

[ErTu93] Erd\H{o}s, Paul and Tuza, Zsolt, Rainbow subgraphs in edge-colorings of complete graphs . (1993), 81--88.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/811}}
\end{document}
