\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#256}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} analysis

\medskip

\noindent\textbf{Problem Statement:}

Let $n\geq 1$ and $f(n)$ be maximal such that for any integers $1\leq a_1\leq \cdots \leq a_n$ we have\[\max_{\lvert z\rvert=1}\left\lvert \prod_{i}(1-z^{a_i})\right\rvert\geq f(n).\]Estimate $f(n)$ - in particular, is it true that there exists some constant $c&#62;0$ such that\[\log f(n) \gg n^c?\]




Erd\H{o}s and Szekeres \cite{ErSz59} proved that $\lim f(n)^{1/n}=1$ and $f(n)&gt;\sqrt{2n}$. Erd\H{o}s proved an upper bound of $\log f(n) \ll n^{1-c}$ for some constant $c&gt;0$ with probabilistic methods. Atkinson \cite{At61} showed that $\log f(n) \ll n^{1/2}\log n$. This was improved to\[\log f(n) \ll n^{1/3}(\log n)^{4/3}\]by Odlyzko \cite{Od82}. If we denote by $f^*(n)$ the analogous quantity with the assumption that $a_1&lt;\cdots&lt;a_n$ then Bourgain and Chang \cite{BoCh18} prove that\[\log f^*(n)\ll (n\log n)^{1/2}\log\log n.\]Atkinson \cite{At61} noted this is related to the Chowla cosine problem [510] , in that if for any set of $n$ integers $A$ there exists $\theta$ such that $\sum_{n\in A}\cos(n\theta) &lt; -M_n$ then\[\log f^*(n) \ll M_n \log n.\]The answer to the specific question asked is no - Belov and Konyagin \cite{BeKo96} proved that\[\log f(n) \ll (\log n)^4.\]




References

[At61] Atkinson, F. V., On a problem of Erd\H{o}s and Szekeres . Canad. Math. Bull. (1961), 7-12.

[BeKo96] Belov, A. S. and Konyagin, S. V., An estimate for the free term of a nonnegative trigonometric polynomial with integer coefficients . Mat. Zametki (1996), 627--629.

[BoCh18] Bourgain, J. and Chang, Mei-Chu, On a paper of Erd\H{o}s and Szekeres . J. Anal. Math. (2018), 253-271.

[ErSz59] Erd\H{o}s, P. and Szekeres, G., On the product $\Pi^n_{k=1}(1-z^ak)$ . Acad. Serbe Sci. Publ. Inst. Math. (1959), 29-34.

[Od82] Odlyzko, A. M., Minima of cosine sums and maxima of polynomials on the unit circle . J. London Math. Soc. (2) (1982), 412-420.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/256}}
\end{document}
