\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#96}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, distances, convex

\medskip

\noindent\textbf{Problem Statement:}

If $n$ points in $\mathbb{R}^2$ form a convex polygon then there are $O(n)$ many pairs which are distance $1$ apart.




Conjectured by Erd\H{o}s and Moser. In \cite{Er92e} Erd\H{o}s credits the conjecture that the true upper bound is $2n$ to himself and Fishburn. F\"{u}redi \cite{Fu90} proved an upper bound of $O(n\log n)$. A short proof of this bound was given by Brass and Pach \cite{BrPa01}. The best known upper bound is\[\leq n\log_2n+4n,\]due to Aggarwal \cite{Ag15}. Edelsbrunner and Hajnal \cite{EdHa91} have constructed $n$ such points with $2n-7$ pairs distance $1$ apart. (This disproved an early stronger conjecture of Erd\H{o}s and Moser, that the true answer was $\frac{5}{3}n+O(1)$.) A positive answer would follow from [97] . See also [90] . In \cite{Er92e} Erd\H{o}s makes the stronger conjecture that, if $g(x)$ counts the largest number of points equidistant from $x$ in $A$, then\[\sum_{x\in A}g(x)&lt; 4n.\]He notes that the example of Edelsbrunner and Hajnal shows that $\sum_{x\in A}g(x)&gt;4n-O(1)$ is possible.




References

[Ag15] Aggarwal, Amol, On unit distances in a convex polygon . Discrete Math. (2015), 88-92.

[BrPa01] Brass , Peter and Pach, J\'{a}nos, The maximum number of times the same distance can occur among the vertices of a convex {$n$}-gon is {$O(n\log n)$} . J. Combin. Theory Ser. A (2001), 178-179.

[EdHa91] Edelsbrunner, Herbert and Hajnal, P\'{e}ter, A lower bound on the number of unit distances between the vertices of a convex polygon . J. Combin. Theory Ser. A (1991), 312-316.

[Er92e] Erd\H{o}s, P\'{a}l, Some Unsolved problems in Geometry, Number Theory and Combinatorics . Eureka (1992), 44-48.

[Fu90] F\"{u}redi, Zolt\'{a}n, The maximum number of unit distances in a convex {$n$}-gon . J. Combin. Theory Ser. A (1990), 316-320.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/96}}
\end{document}
