\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#529}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, probability

\medskip

\noindent\textbf{Problem Statement:}

Let $d_k(n)$ be the expected distance from the origin after taking $n$ random steps from the origin in $\mathbb{Z}^k$ (conditional on no self intersections) - that is, a self-avoiding walk . Is it true that\[\lim_{n\to \infty}\frac{d_2(n)}{n^{1/2}}= \infty?\]Is it true that\[d_k(n)\ll n^{1/2}\]for $k\geq 3$?




Slade \cite{Sl87} proved that, for $k$ sufficiently large, $d_k(n)\sim Dn^{1/2}$ for some constant $D&gt;0$ (independent of $k$). Hara and Slade (\cite{HaSl91} and \cite{HaSl92}) proved this for all $k\geq 5$. For $k=2$ Duminil-Copin and Hammond \cite{DuHa13} have proved that $d_2(n)=o(n)$. It is now conjectured that $d_k(n)\ll n^{1/2}$ is false for $k=3$ and $k=4$, and more precisely (see for example Section 1.4 of \cite{MaSl93}) that $d_2(n)\sim Dn^{3/4}$, $d_3(n)\sim n^{\nu}$ where $\nu\approx 0.59$, and $d_4(n)\sim D(\log n)^{1/8}n^{1/2}$. Madras and Slade \cite{MaSl93} have a monograph on the topic of self-avoiding walks. See also [528] .




References

[DuHa13] Duminil-Copin, Hugo and Hammond, Alan, Self-avoiding walk is sub-ballistic . Comm. Math. Phys. (2013), 401--423.

[HaSl91] Hara, Takashi and Slade, Gordon, Critical behaviour of self-avoiding walk in five or more dimensions . Bull. Amer. Math. Soc. (N.S.) (1991), 417--423.

[HaSl92] Hara, Takashi and Slade, Gordon, Self-avoiding walk in five or more dimensions. I. The critical behaviour . Comm. Math. Phys. (1992), 101--136.

[MaSl93] Madras, Neal and Slade, Gordon, The self-avoiding walk . (1993), xiv+425.

[Sl87] Slade, Gordon, The diffusion of self-avoiding random walk in high dimensions . Comm. Math. Phys. (1987), 661--683.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/529}}
\end{document}
