\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#169}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} additive combinatorics, arithmetic progressions

\medskip

\noindent\textbf{Problem Statement:}

Let $k\geq 3$ and $f(k)$ be the supremum of $\sum_{n\in A}\frac{1}{n}$ as $A$ ranges over all sets of positive integers which do not contain a $k$-term arithmetic progression. Estimate $f(k)$. Is\[\lim_{k\to \infty}\frac{f(k)}{\log W(k)}=\infty\]where $W(k)$ is the van der Waerden number?




Berlekamp \cite{Be68} proved $f(k) \geq \frac{\log 2}{2}k$. Gerver \cite{Ge77} proved\[f(k) \geq (1-o(1))k\log k.\]It is trivial that\[\frac{f(k)}{\log W(k)}\geq \frac{1}{2},\]but improving the right-hand side to any constant $&gt;1/2$ is open. Gerver also proved (see the comments for an alternative argument of Tao) that [3] is equivalent to $f(k)$ being finite for all $k$. The current record for $f(3)$ is $f(3)\geq 3.00849$, due to Wr\'{o}blewski \cite{Wr84}. Walker \cite{Wa25} proved $f(4)\geq 4.43975$. Walker \cite{Wa25} has shown that it suffices to consider Kempner sets (that is, sets of integers defined as all those whose base $b$ digits are contained in some $S\subset \{0,\ldots,b-1\}$ for fixed $b$ and $S$), in the sense that for any $k\geq 3$ and $\epsilon&gt;0$ there is a Kempner set $A$ lacking $k$-term arithmetic progressions such that\[\sum_{n\in A}\frac{1}{n}\geq f(k)-\epsilon.\]In \cite{Er80} Erd\H{o}s asks whether for every $\epsilon&gt;0$ and $k\geq 3$ if $A$ is a set of integers without a $k$-term arithmetic progression such that $\min(A)$ is sufficiently large in terms of $\epsilon$ and $k$ then $\sum_{n\in A}\frac{1}{n}&lt;\epsilon$.




References

[Be68] Berlekamp, E. R., A construction for partitions which avoid long arithmetic progressions . Canad. Math. Bull. (1968), 409-414.

[Er80] Erd\H{o}s, Paul, A survey of problems in combinatorial number theory . Ann. Discrete Math. (1980), 89-115.

[Ge77] Gerver, Joseph L., The sum of the reciprocals of a set of integers with no arithmetic progression of {$k$} terms . Proc. Amer. Math. Soc. (1977), 211--214.

[Wa25] A. Walker, Integer sets of large harmonic sum which avoid long arithmetic progressions . arXiv:2203.06045 (2025).

[Wr84] Wr\'oblewski, J., A nonaveraging set of integers with a large sum of reciprocals . Math. Comp. (1984), 261--262.

\noindent\textbf{Related OEIS sequences:} A005346


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/169}}
\end{document}
