\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#561}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} graph theory, ramsey theory

\medskip

\noindent\textbf{Problem Statement:}

Let $\hat{R}(G)$ denote the size Ramsey number, the minimal number of edges $m$ such that there is a graph $H$ with $m$ edges such that in any $2$-colouring of the edges of $H$ there is a monochromatic copy of $G$. Let $F_1$ and $F_2$ be the union of stars. More precisely, let $F_1=\cup_{i\leq s} K_{1,n_i}$ and $F_2=\cup_{j\leq t} K_{1,m_j}$ with $n_1\geq \cdots \geq n_s\geq 1$ and $m_1\geq \cdots \geq m_t\geq 1$. Prove that\[\hat{R}(F_1,F_2) = \sum_{2\leq k\leq s+t}l_k\]where\[l_k=\max\{n_i+m_j-1 : i+j=k\}.\]




Burr, Erd\H{o}s, Faudree, Rousseau, and Schelp \cite{BEFRS78} proved this when all the $n_i$ are identical and all the $m_i$ are identical. Gy\H{o}ri and Schelp \cite{GySc02} proved this when $\binom{l_k}{2}&gt; \sum_{k\leq i\leq s+t}l_i$ for all $2\leq k\leq s+t$. More special cases, including all cases when $s=1$, or $s=2$ and $n_1=n_2$, or all $n_i$ and $m_j$ are odd, or all $n_i$ are an identical odd number and $m_1$ is odd, were proved by Davoodi, Javadi, Kamranian, and Raeisi \cite{DJKR25}. This problem is #30 in Ramsey Theory in the graphs problem collection.




References

[BEFRS78] Burr, S. A. and Erd\H{o}s, P. and Faudree, R. J. and Rousseau, C. C. and Schelp, R. H., Ramsey-minimal graphs for multiple copies . Nederl. Akad. Wetensch. Indag. Math. (1978), 187-195.

[DJKR25] Davoodi, Akbar and Javadi, Ramin and Kamranian, Azam and Raeisi, Ghaffar, On a conjecture of {E}rd\H{o}s on size {R}amsey number of star forests . Ars Math. Contemp. (2025), Paper No. 9, 10.

[GySc02] Gy\H ori, Ervin and Schelp, R. H., Two-edge colorings of graphs with bounded degree in both colors . Discrete Math. (2002), 105--110.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/561}}
\end{document}
