\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#687}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{Prize: \$1000}

\noindent\textbf{Tags:} number theory

\medskip

\noindent\textbf{Problem Statement:}

Let $Y(x)$ be the maximal $y$ such that there exists a choice of congruence classes $a_p$ for all primes $p\leq x$ such that every integer in $[1,y]$ is congruent to at least one of the $a_p\pmod{p}$. Give good estimates for $Y(x)$. In particular, can one prove that $Y(x)=o(x^2)$ or even $Y(x)\ll x^{1+o(1)}$?




This function (associated with Jacobsthal) is closely related to the problem of gaps between primes (see [4] ). The best known upper bound is due to Iwaniec \cite{Iw78},\[Y(x) \ll x^2.\]The best lower bound is due to Ford, Green, Konyagin, Maynard, and Tao \cite{FGKMT18},\[Y(x) \gg x\frac{\log x\log\log\log x}{\log\log x},\]improving on a previous bound of Rankin \cite{Ra38}. Maier and Pomerance have conjectured that $Y(x)\ll x(\log x)^{2+o(1)}$. In \cite{Er80} he writes 'It is not clear who first formulated this problem - probably many of us did it independently. I offer the maximum of \$1000 dollars and $1/2$ my total savings for clearing up of this problem.' In \cite{Er80} Erd\H{o}s also asks about a weaker variant in which all except $o(y/\log y)$ of the integers in $[1,y]$ are congruent to at least one of the $a_p\pmod{p}$, and in particular asks if the answer is very different. See also [688] and [689] . A more general Jacobsthal function is the focus of [970] .




References

[Er80] Erd\H{o}s, Paul, A survey of problems in combinatorial number theory . Ann. Discrete Math. (1980), 89-115.

[FGKMT18] Ford, Kevin and Green, Ben and Konyagin, Sergei and Maynard, James and Tao, Terence, Long gaps between primes . J. Amer. Math. Soc. (2018), 65-105.

[Iw78] Iwaniec, Henryk, On the problem of {J}acobsthal . Demonstratio Math. (1978), 225--231.

[Ra38] Rankin, R. A., The Difference between Consecutive Prime Numbers . J. London Math. Soc. (1938), 242-247.

\noindent\textbf{Related OEIS sequences:} A048670, A058989


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/687}}
\end{document}
