\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#571}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} graph theory, turan number

\medskip

\noindent\textbf{Problem Statement:}

Show that for any rational $\alpha \in [1,2)$ there exists a bipartite graph $G$ such that\[\mathrm{ex}(n;G)\asymp n^{\alpha}.\]




A problem of Erd\H{o}s and Simonovits. In \cite{Er78} Erd\H{o}s wrote 'I am not entirely sure that a trivial counterexample can not be found'. Bukh and Conlon \cite{BuCo18} proved that this holds if we weaken asking for the extremal number of a single graph to asking for the extremal number of a finite family of graphs. A rational $\alpha\in [1,2)$ for which this holds is known as a Tur\'{a}n exponent. Known Tur\'{a}n exponents are: {UL} {LI} $\frac{3}{2}-\frac{1}{2s}$ for $s\geq 2$ (Conlon, Janzer, and Lee \cite{CJL21}).{/LI} {LI} $\frac{4}{3}-\frac{1}{3s}$ and $\frac{5}{4}-\frac{1}{4s}$ for $s\geq 2$ (Jiang and Qiu \cite{JiQi20}).{/LI} {LI} $2-\frac{a}{b}$ for $\lfloor b/a\rfloor^3 \leq a\leq \frac{b}{\lfloor b/a\rfloor+1}+1$ (Jiang, Jiang, and Ma \cite{JJM20}).{/LI} {LI} $2-\frac{a}{b}$ with $b&gt;a\geq 1$ and $b\equiv \pm 1\pmod{a}$ (Kang, Kim, and Liu \cite{KKL21}).{/LI} {LI} $1+a/b$ with $b&gt;a^2$ (Jiang and Qiu \cite{JiQi23}),{/LI} {LI} $2-\frac{2}{2b+1}$ for $b\geq 2$ or $7/5$ (Jiang, Ma, and Yepremyan \cite{JMY22}).{/LI} {LI} $2-a/b$ with $b\geq (a-1)^2$ (Conlon and Janzer \cite{CoJa22}).{/LI} {/UL} See also [713] . This problem is #45 in Extremal Graph Theory in the graphs problem collection.




References

[BuCo18] Bukh, Boris and Conlon, David, Rational exponents in extremal graph theory . J. Eur. Math. Soc. (JEMS) (2018), 1747-1757.

[CJL21] Conlon, David and Janzer, Oliver and Lee, Joonkyung, More on the extremal number of subdivisions . Combinatorica (2021), 465-494.

[CoJa22] Conlon, David and Janzer, Oliver, Rational exponents near two . Adv. Comb. (2022), Paper No. 9, 10.

[Er78] Erd\H{o}s, Paul, Problems and results in combinatorial analysis and combinatorial number theory . Proceedings of the Ninth Southeastern Conference on Combinatorics, Graph Theory, and Computing (Florida Atlantic Univ., Boca Raton, Fla., 1978) (1978), 29-40.

[JJM20] Jiang, Tao and Jiang, Zilin and Ma, Jie, Negligible obstructions and Tur\'{a}n exponents . arXiv:2007.02975 (2020).

[JMY22] Jiang, Tao and Ma, Jie and Yepremyan, Liana, On Tur\'{a}n exponents of bipartite graphs . Combin. Probab. Comput. (2022), 333-344.

[JiQi20] Jiang, Tao and Qiu, Yu, Tur\'{a}n numbers of bipartite subdivisions . SIAM J. Discrete Math. (2020), 556-570.

[JiQi23] Jiang, Tao and Qiu, Yu, Many Tur\'{a}n exponents via subdivisions . Combin. Probab. Comput. (2023), 134-150.

[KKL21] Kang, Dong Yeap and Kim, Jaehoon and Liu, Hong, On the rational Tur\'{a}n exponents conjecture . J. Combin. Theory Ser. B (2021), 149-172.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/571}}
\end{document}
