\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#671}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{Prize: \$250}

\noindent\textbf{Tags:} analysis

\medskip
\noindent\textbf{Problem Statement:}

Given $a_{i}^n\in [-1,1]$ for all $1\leq i\leq n<\infty$ we define $p_{i}^n$ as the unique polynomial of degree $n-1$ such that $p_{i}^n(a_{i}^n)=1$ and $p_{i}^n(a_{i'}^n)=0$ if $1\leq i'\leq n$ with $i\neq i'$. We similarly define\[\mathcal{L}^nf(x) = \sum_{1\leq i\leq n}f(a_i^n)p_i^n(x),\]the unique polynomial of degree $n-1$ which agrees with $f$ on $a_i^n$ for $1\leq i\leq n$ (that is, the sequence of Lagrange interpolation polynomials).

Is there such a sequence of $a_i^n$ such that for every continuous $f:[-1,1]\to \mathbb{R}$ there exists some $x\in [-1,1]$ where\[\limsup_{n\to \infty} \sum_{1\leq i\leq n}\lvert p_{i}^n(x)\rvert=\infty\]and yet\[\mathcal{L}^nf(x) \to f(x)?\]Is there such a sequence such that\[\limsup_{n\to \infty} \sum_{1\leq i\leq n}\lvert p_{i}^n(x)\rvert=\infty\]for every $x\in [-1,1]$ and yet for every continuous $f:[-1,1]\to \mathbb{R}$ there exists $x\in [-1,1]$ with\[\mathcal{L}^nf(x) \to f(x)?\]

\bigskip
\noindent\textbf{Remarks:}

Bernstein \cite{Be31} proved that for any choice of $a_i^n$ there exists $x_0\in [-1,1]$ such that\[\limsup_{n\to \infty} \sum_{1\leq i\leq n}\lvert p_{i}^n(x)\rvert=\infty.\]Erd\H{o}s and V\'{e}rtesi \cite{ErVe80} proved that for any choice of $a_i^n$ there exists a continuous $f:[-1,1]\to \mathbb{R}$ such that\[\limsup_{n\to \infty} \lvert \mathcal{L}^nf(x)\rvert=\infty\]for almost all $x\in [-1,1]$.

\bigskip
\noindent\textbf{References:}

[Be31] S. Bernstein, Sur la limitation des valeurs d'un polynome $P_n(x)$ de degr\'{e} n sur tout un segment par ses valeurs en $(n+1)$ points du segment. Izv. Akad. Nauk. SSSR (1931), 1025-1050.
 
 [ErVe80] Erd\H{o}s, P. and V\'{e}rtesi, P., On the almost everywhere divergence of Lagrange
interpolatory polynomials for arbitrary system of nodes. Acta Math. Acad. Sci. Hungar. (1980), 71-89.

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/671}}
\end{document}
