\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#212}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} geometry, distances

\medskip

\noindent\textbf{Problem Statement:}

Is there a dense subset of $\mathbb{R}^2$ such that all pairwise distances are rational?




Conjectured by Ulam. Erd\H{o}s believed there cannot be such a set. This problem is discussed in a blogpost by Terence Tao, in which he shows that there cannot be such a set, assuming the Bombieri-Lang conjecture. The same conclusion was independently obtained by Shaffaf \cite{Sh18}. Indeed, Shaffaf and Tao actually proved that such a rational distance set must be contained in a finite union of real algebraic curves. Solymosi and de Zeeuw \cite{SdZ10} then proved (unconditionally) that a rational distance set contained in a real algebraic curve must be finite, unless the curve contains a line or a circle. Ascher, Braune, and Turchet \cite{ABT20} observed that, combined, these facts imply that a rational distance set in general position must be finite (conditional on the Bombieri-Lang conjecture). In \cite{Er87b} Erd\H{o}s mentions that Besicovitch conjectured that the limit points of a rational distance set cannot contain arbitrarily large convex sets.




References

[ABT20] Ascher, K. and Braune, L. and Turchet, A., The Erd\H{o}s-Ulam problem, Lang's conjecture, and uniformity . arXiv:1901.02616 (2020).

[Er87b] Erd\H{o}s, P., Some combinatorial and metric problems in geometry . Intuitive geometry (Si\'{o}fok, 1985) (1987), 167-177.

[SdZ10] Solymosi, Jozsef and de Zeeuw, Frank, On a question of Erd\H{o}s and Ulam . Discrete Comput. Geom. (2010), 393-401.

[Sh18] Shaffaf, Jafar, A solution of the Erd\H{o}s-Ulam problem on rational distance sets assuming the Bombieri-Lang conjecture . Discrete Comput. Geom. (2018), 283-293.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/212}}
\end{document}
