\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#30}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{Prize: \$1000}

\noindent\textbf{Tags:} number theory, sidon sets, additive combinatorics

\medskip

\noindent\textbf{Problem Statement:}

Let $h(N)$ be the maximum size of a Sidon set in $\{1,\ldots,N\}$. Is it true that, for every $\epsilon&#62;0$,\[h(N) = N^{1/2}+O_\epsilon(N^\epsilon)?\]




A problem of Erd\H{o}s and Tur\'{a}n. It may even be true that $h(N)=N^{1/2}+O(1)$, but Erd\H{o}s remarks this is perhaps too optimistic. Erd\H{o}s and Tur\'{a}n \cite{ErTu41} proved an upper bound of $N^{1/2}+O(N^{1/4})$, with an alternative proof by Lindstr\"{o}m \cite{Li69}. Both proofs in fact give\[h(N) \leq N^{1/2}+N^{1/4}+1.\]Balogh, F\"{u}redi, and Roy \cite{BFR21} improved the bound in the error term to $0.998N^{1/4}$. This was further optimised by O'Bryant \cite{OB22}. The current record is\[h(N)\leq N^{1/2}+0.98183N^{1/4}+O(1),\]due to Carter, Hunter, and O'Bryant \cite{CHO25}. Singer \cite{Si38} was the first to show that $h(N)\geq (1-o(1))N^{1/2}$ for all $N$. For a detailed survey of the literature we refer to \cite{OB04}. See also [241] and [840] . This problem is Problem 31 on Green's open problems list . This is discussed in problem C9 of Guy's collection \cite{Gu04}.




References

[BFR21] Balogh, J. and F\"{u}redi, Z. and Roy, S., An upper bound on the size of Sidon sets . arXiv:2103.15850 (2021).

[CHO25] Carter, D. and Hunter, Z. and O'Bryant, K., On the diameter of finite Sidon sets . Acta Math. Hungar. (2025), 108--126.

[ErTu41] Erd\H{o}s, P. and Tur\'{a}n, P., On a problem of Sidon in additive number theory, and on some related problems . J. London Math. Soc. (1941), 212-215.

[Gu04] Guy, Richard K., Unsolved problems in number theory . (2004), xviii+437.

[Li69] Lindstr\"{o}m, B., An inequality for $B_2$-sequences . J. Combinatorial Theory (1969), 211-212.

[OB04] O'Bryant, Kevin, A complete annotated bibliography of work related to Sidon sequences . Electron. J. Combin. (2004), 39.

[OB22] O'Bryant, K., On the size of finite Sidon sets . arXiv:2207.07800 (2022).

[Si38] Singer, James, A theorem in finite projective geometry and some applications to number theory . Trans. Amer. Math. Soc. (1938), 377--385.

\noindent\textbf{Related OEIS sequences:} A143824, A227590, A003022


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/30}}
\end{document}
