\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1206}
\date{Last updated: 2026-04-04}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, sidon sets

\medskip

\noindent\textbf{Problem Statement:}

Does $\{1,2^3,\ldots,N^3\}$ contain a Sidon set of size $\gg N$? Is there an infinite set $A\subset \mathbb{N}$ of positive density such that $\{a^3 : a\in A\}$ is a Sidon set?




This question may also be asked for higher powers, and [324] suggests that for fifth powers $\{n^5 :n\geq 1\}$ may itself be a Sidon set. Gabdullin and Konyagin \cite{GaKo24} proved that there is a constant $c&gt;0$ such that $\{ n^3 : N-cN^{1/2}\leq n\leq N\}$ is a Sidon set. Garaev, Garayev, and Konyagin \cite{GGK26} have proved that the exponent $1/2$ can be improved to $4/7-o(1)$ for infinitely many $N$, and to $3/5$ for all $N$ if we replace $n^3$ with $n^4$. In \cite{Er80} Erd\H{o}s defines $g_k(A)$ to be the maximal size of a Sidon set of $\{a^k : a\in A\}$, and asks whether $g_k(A)\geq g_k(\{1,\ldots,N\})$ where $N=\lvert A\rvert$. (See [530] for the linear analogue.) See also [773] for an analogous question about squares.




References

[Er80] Erd\H{o}s, Paul, A survey of problems in combinatorial number theory . Ann. Discrete Math. (1980), 89-115.

[GGK26] M. Garaev, F. Garayev, and S. Konyagin, On Sidon sets with squares, cubes, and quartics in short intervals . arXiv:2602.08807 (2026).

[GaKo24] Gabdullin, M. R. and Konyagin, S. V., Trigonometric polynomials with frequencies in the set of cubes . Math. Notes (2024), 336--340.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1206}}
\end{document}
