\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#975}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, divisors

\medskip

\noindent\textbf{Problem Statement:}

Let $f\in \mathbb{Z}[x]$ be an irreducible non-constant polynomial such that $f(n)\geq 1$ for all large $n\in\mathbb{N}$. Does there exist a constant $c=c(f)&#62;0$ such that\[\sum_{n\leq X} \tau(f(n))\sim cX\log X,\]where $\tau$ is the divisor function?




Van der Corput \cite{Va39} proved that\[\sum_{n\leq X} \tau(f(n))\gg_f X\log X.\]Erd\H{o}s \cite{Er52b} proved using elementary methods that\[\sum_{n\leq X} \tau(f(n))\ll_f X\log X.\]Such an asymptotic formula is known whenever $f$ is an irreducible quadratic, as proved by Hooley \cite{Ho63}. The form of $c$ depends on $f$ in a complicated fashion (see the work of McKee \cite{Mc95}, \cite{Mc97}, and \cite{Mc99} for expressions for various types of quadratic $f$). For example,\[\sum_{n\leq x}\tau(n^2+1)=\frac{3}{\pi}x\log x+O(x).\]Tao has a blog post on this topic.




References

[Er52b] Erd\H{o}s, P., On the sum {$\sum^x_{k=1} d(f(k))$} . J. London Math. Soc. (1952), 7--15.

[Ho63] Hooley, Christopher, On the number of divisors of a quadratic polynomial . Acta Math. (1963), 97--114.

[Mc95] McKee, James, On the average number of divisors of quadratic polynomials . Math. Proc. Cambridge Philos. Soc. (1995), 389--392.

[Mc97] McKee, James, A note on the number of divisors of quadratic polynomials . (1997), 275--281.

[Mc99] McKee, James, The average number of divisors of an irreducible quadratic polynomial . Math. Proc. Cambridge Philos. Soc. (1999), 17--22.

[Va39] van der Corput, J. G., Une in\'{e}galit\'{e}{} relative au nombre des diviseurs . Nederl. Akad. Wetensch., Proc. (1939), 547--553.

\noindent\textbf{Related OEIS sequences:} A147807, possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/975}}
\end{document}
