\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#787}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} additive combinatorics

\medskip

\noindent\textbf{Problem Statement:}

Let $g(n)$ be maximal such that given any set $A\subset \mathbb{R}$ with $\lvert A\rvert=n$ there exists some $B\subseteq A$ of size $\lvert B\rvert\geq g(n)$ such that $b_1+b_2\not\in A$ for all $b_1\neq b_2\in B$. Estimate $g(n)$.




This function was considered by Erd\H{o}s and Moser. Choi observed that, without loss of generality, one can assume that $A\subset \mathbb{Z}$. Klarner proved $g(n) \gg \log n$ (indeed, a greedy construction suffices). Choi \cite{Ch71} proved $g(n) \ll n^{2/5+o(1)}$. The current best bounds known are\[(\log n)^{1+c} \ll g(n) \ll \exp(\sqrt{\log n})\]for some constant $c&#62;0$, the lower bound due to Sanders \cite{Sa21} and the upper bound due to Ruzsa \cite{Ru05}. Beker \cite{Be25} has proved\[(\log n)^{1+\tfrac{1}{68}+o(1)} \ll g(n).\]




References

[Be25] A. Beker, The Erd\H{o}s-Moser sum-free set problem via improved bounds for $k$-configurations . arXiv:2501.10203 (2025).

[Ch71] Choi, S. L. G., On a combinatorial problem in number theory . Proc. London Math. Soc. (3) (1971), 629-642.

[Ru05] Ruzsa, Imre Z., Sum-avoiding subsets . Ramanujan J. (2005), 77-82.

[Sa21] Sanders, Tom, The Erd\H{o}s-Moser sum-free set problem . Canad. J. Math. (2021), 63-107.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/787}}
\end{document}
