\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#86}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{Prize: \$100}

\noindent\textbf{Tags:} graph theory

\medskip

\noindent\textbf{Problem Statement:}

Let $Q_n$ be the $n$-dimensional hypercube graph (so that $Q_n$ has $2^n$ vertices and $n2^{n-1}$ edges). Is it true that every subgraph of $Q_n$ with\[\geq \left(\frac{1}{2}+o(1)\right)n2^{n-1}\]many edges contains a $C_4$?




Let $f(n)$ be the maximum number of edges in a subgraph of $Q_n$ without a $C_4$, so that this conjecture is that $f(n)\leq (\frac{1}{2}+o(1))n2^{n-1}$. Erd\H{o}s \cite{Er91} showed that\[f(n) \geq \left(\frac{1}{2}+\frac{c}{n}\right)n2^{n-1}\]for some constant $c&gt;0$, and wrote it is 'perhaps not hopeless' to determine $f(n)$ exactly. Brass, Harborth, and Nienborg \cite{BHN95} improved this to\[f(n) \geq \left(\frac{1}{2}+\frac{c}{\sqrt{n}}\right)n2^{n-1}\]for some constant $c&gt;0$. Balogh, Hu, Lidicky, and Liu \cite{BHLL14} proved that $f(n)\leq 0.6068 n2^{n-1}$. This was improved to $\leq 0.60318 n2^{n-1}$ by Baber \cite{Ba12b}. A similar question can be asked for other even cycles. See also [666] and the entry in the graphs problem collection .




References

[BHLL14] Balogh, J\'{o}zsef and Hu, Ping and Lidick\'{y}, Bernard and Liu, Hong, Upper bounds on the size of 4- and 6-cycle-free subgraphs of the hypercube . European J. Combin. (2014), 75-85.

[BHN95] Brass, Peter and Harborth, Heiko and Nienborg, Hauke, On the maximum number of edges in a {$C_4$}-free subgraph of {$Q_n$} . J. Graph Theory (1995), 17--23.

[Ba12b] R. Baber, Tur\'{a}n densities of hypercubes . arXiv:1201.3587 (2012).

[Er91] Erd\H{o}s, P., Problems and results in combinatorial analysis and combinatorial number theory . Graph theory, combinatorics, and applications, Vol. 1 (Kalamazoo, MI, 1988) (1991), 397-406.

\noindent\textbf{Related OEIS sequences:} A245762


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/86}}
\end{document}
