\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#996}
\date{Last updated: 2025-09-07}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} analysis

\medskip
\noindent\textbf{Problem Statement:}

Let $n_1<n_2<\cdots$ be a lacunary sequence of integers, and let $f\in L^2([0,1])$. Let $f_n$ be the $n$th partial sum of the Fourier series of $f(x)$. Is there an absolute constant $C>0$ such that, if\[\| f-f_n\|_2 \ll \frac{1}{(\log\log\log n)^{C}}\]then\[\lim_{N\to\infty}\frac{1}{N}\sum_{k\leq N}f(\{\alpha n_k\})=\int_0^1 f(x)\mathrm{d}x\]for almost every $\alpha$?

\bigskip
\noindent\textbf{Remarks:}

Raikov proved the conclusion always holds (for every $f\in L^2([0,1])$, with no assumption on $\| f-f_n\|_2$) if $n_k=a^k$ for some integer $a\geq 2$. Erd\H{o}s \cite{Er64b} also asked whether this is true for $n_k=\lfloor a^k\rfloor$ for some $a>1$.

Kac, Salem, and Zygmund \cite{KSZ48} proved that the conclusion holds if\[\| f-f_n\|_2 \ll \frac{1}{(\log n)^{c}}\]for some constant $c>1$. Erd\H{o}s \cite{Er49d} proved that the conclusion holds if\[\| f-f_n\|_2 \ll \frac{1}{(\log\log n)^{c}}\]for some constant $c>1$. Matsuyama \cite{Ma66} improved this to $c>1/2$.

In \cite{Er64b} Erd\H{o}s asked whether the conclusion holds for all bounded functions $f$ and lacunary sequences $n_k$.

\bigskip
\noindent\textbf{References:}

[Er49d] Erd\H{o}s, P., On the strong law of large numbers. Trans. Amer. Math. Soc. (1949), 51--56.
 
 [Er64b] Erd\H{o}s, P., Problems and results on diophantine approximations. Compositio Math. (1964), 52-65.
 
 [KSZ48] Kac, M. and Salem, R. and Zygmund, A., A gap theorem. Trans. Amer. Math. Soc. (1948), 235--243.
 
 [Ma66] Matsuyama, Noboru, On the strong law of large numbers. Tohoku Math. J. (2) (1966), 259--269.

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/996}}
\end{document}
