\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#5}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, primes

\medskip

\noindent\textbf{Problem Statement:}

Let $C\geq 0$. Is there an infinite sequence of $n_i$ such that\[\lim_{i\to \infty}\frac{p_{n_i+1}-p_{n_i}}{\log n_i}=C?\]




Let $S$ be the set of limit points of $(p_{n+1}-p_n)/\log n$. This problem asks whether $S=[0,\infty]$. Although this conjecture remains unproven, a lot is known about $S$. Some highlights: {UL} {LI}$\infty\in S$ by Westzynthius' result \cite{We31} on large prime gaps,{/LI} {LI}$0\in S$ by the work of Goldston, Pintz, and Yildirim \cite{GPY09} on small prime gaps,{/LI} {LI}Erd\H{o}s \cite{Er55} and Ricci \cite{Ri56} independently showed that $S$ has positive Lebesgue measure,{/LI} {LI} Hildebrand and Maier \cite{HiMa88} showed that $S$ contains arbitrarily large (finite) numbers,{/LI} {LI} Pintz \cite{Pi16} showed that there exists some small constant $c&gt;0$ such that $[0,c]\subset S$,{/LI} {LI} Banks, Freiberg, and Maynard \cite{BFM16} showed that at least $12.5\%$ of $[0,\infty)$ belongs to $S$,{/LI} {LI} Merikoski \cite{Me20} showed that at least $1/3$ of $[0,\infty)$ belongs to $S$, and that $S$ has bounded gaps.{/LI} {/UL} In \cite{Er65b}, \cite{Er85c}, and \cite{Er97c} Erd\H{o}s asks whether $S$ is everywhere dense (but Weisenberg notes that clearly $S$ is closed so this is equivalent to asking whether $S=[0,\infty]$). See also [234] .




References

[BFM16] Banks, William D. and Freiberg, Tristan and Maynard, James, On limit points of the sequence of normalized prime gaps . Proc. Lond. Math. Soc. (3) (2016), 515-539.

[Er55] Erd\H{o}s, Paul, Some remarks on number theory . Riveon Lematematika (1955), 45-48.

[Er65b] Erd\H{o}s, Paul, Some recent advances and current problems in number theory . Lectures on Modern Mathematics, Vol. III (1965), 196-244.

[Er85c] Erd\H{o}s, P., On some of my problems in number theory I would most like to see solved . Number theory (Ootacamund, 1984) (1985), 74-84.

[Er97c] Erd\H{o}s, Paul, Some of my favorite problems and results . The mathematics of Paul Erd\H{o}s, I (1997), 47-67.

[GPY09] Goldston, Daniel A. and Pintz, J\'{a}nos and Y\i ld\i r\i m, Cem Y., Primes in tuples. I . Ann. of Math. (2) (2009), 819-862.

[HiMa88] Hildebrand, Adolf and Maier, Helmut, Gaps between prime numbers . Proc. Amer. Math. Soc. (1988), 1-9.

[Me20] Merikoski, Jori, Limit points of normalized prime gaps . J. Lond. Math. Soc. (2) (2020), 99-124.

[Pi16] Pintz, J\'{a}nos, Polignac numbers, conjectures of Erd\H{o}s on gaps between primes, arithmetic progressions in primes, and the bounded gap conjecture . From arithmetic to zeta-functions (2016), 367-384.

[Ri56] Ricci, Giovanni, Recherches sur l'allure de la suite $\{p_{n+1}-p_n/\log p_n\}$ . Colloque sur la Th\'{e}orie des Nombres, Bruxelles, 1955 (1956), 93-106.

[We31] Westzynthius, E., \"{U}ber die Verteilung der Zahlen, die zu den n ersten Primzahlen teilerfremd sind . Commentat. Phys. Math. (1931), 1-37.

\noindent\textbf{Related OEIS sequences:} A001223


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/5}}
\end{document}
