\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1095}
\date{Last updated: 2025-10-18}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, binomial coefficients

\medskip

\noindent\textbf{Problem Statement:}

Let $g(k)&#62;k+1$ be the smallest $n$ such that all prime factors of $\binom{n}{k}$ are $&#62;k$. Estimate $g(k)$.




A question of Ecklund, Erd\H{o}s, and Selfridge \cite{EES74}, who proved\[k^{1+c}&lt;g(k)\leq \exp((1+o(1))k)\]for some constant $c&gt;0$, and conjectured $g(k)&lt;L_k=[1,\ldots,k]$, the least common multiple of all integers $\leq k$, for all large $k$. In \cite{EES74} they further conjecture that\[\limsup \frac{g(k+1)}{g(k)}=\infty\]and\[\liminf \frac{g(k+1)}{g(k)}=0.\]The lower bound was improved by Erd\H{o}s, Lacampagne, and Selfridge \cite{ELS93} and Granville and Ramar\'{e} \cite{GrRa96}. The current record is\[g(k) \gg \exp(c(\log k)^2)\]for some $c&gt;0$, due to Konyagin \cite{Ko99b}. Erd\H{o}s, Lacampagne, and Selfridge \cite{ELS93} write 'it is clear to every right-thinking person' that $g(k)\geq\exp(c\frac{k}{\log k})$ for some constant $c&gt;0$. Sorenson, Sorenson, and Webster \cite{SSW20} give heuristic evidence that\[\log g(k) \asymp \frac{k}{\log k}.\]See also [1094] .




References

[EES74] Ecklund, Jr., E. F. and Erd\H{o}s, P. and Selfridge, J. L., A new function associated with the prime factors of {$(\sp{n}\sb{k})$} . Math. Comp. (1974), 647--649.

[ELS93] Erd\H{o}s, P. and Lacampagne, C. B. and Selfridge, J. L., Estimates of the least prime factor of a binomial coefficient . Math. Comp. (1993), 215--224.

[GrRa96] Granville, Andrew and Ramar\'{e}, Olivier, Explicit bounds on exponential sums and the scarcity of squarefree binomial coefficients . Mathematika (1996), 73--107.

[Ko99b] Konyagin, S. V., Estimates of the least prime factor of a binomial coefficient . Mathematika (1999), 41--55.

[SSW20] Sorenson, Brianna and Sorenson, Jonathan and Webster, Jonathan, An algorithm and estimates for the {E}rd\H{o}s-Selfridge function . (2020), 371--385.

\noindent\textbf{Related OEIS sequences:} A003458


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1095}}
\end{document}
