\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1188}
\date{Last updated: 2026-04-04}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, covering systems

\medskip

\noindent\textbf{Problem Statement:}

Call a set of distinct integers $1&#60;n_1&#60;\cdots&#60;n_k$ with associated congruence classes $a_i\pmod{n_i}$ a distinct covering system if every integer satisfies at least one of these congruences. A minimal distinct covering system is one such that no proper subset forms a covering system. Let $F(x)$ count the number of minimal distinct covering systems with all moduli in $[1,x]$. Estimate $F(x)$.




Erd\H{o}s \cite{Er80} asked this without the minimal assumption, but otherwise this is trivial (since e.g. any system containing some fixed covering system would work, resulting in $\gg 2^x$ many such systems). He expected $F(x)\to \infty$ as $x\to \infty$ 'very slowly'. In \cite{Er80} Erd\H{o}s also asked to estimate the number of irreducible covering sets in $[1,x]$ (that is, those sets of moduli which contain no proper sets of moduli which form a covering system, perhaps with a different choice of congruence class). If this count is $G(x)$ then trivially $G(x)\leq F(x)$. See [1189] for more on irreducible covering sets. It follows from the solution of Hough \cite{Ho15} to [2] that $F(x)\to \infty$ as $x\to \infty$ (although a lower bound of $F(x) \gg \log x$ is easily established by elementary means, as observed by van Doorn in the comments). The lower bound construction of Balister, Bollob\'{a}s, Morris, Sahasrabudhe, and Tiba \cite{BBMST24} implies that there are at least $\exp(k^{3+o(1)})$ many minimal distinct covering systems, all of whose moduli are at most $\prod_{p&lt;k}p=e^{(1+o(1))k}$, which implies\[F(x)\geq \exp((\log x)^{3-o(1)})\](in particular $F(x)$ grows faster than any polynomial, perhaps contradicting Erd\H{o}s' guess that it should grow 'very slowly', although it is possible he was thinking of the number of different moduli possible). The trivial upper bound is\[F(x) \leq \exp(O(x\log x)).\]




References

[BBMST24] Balister, Paul and Bollob\'{a}s, B\'ela and Morris, Robert and Sahasrabudhe, Julian and Tiba, Marius, The structure and number of {E}rd\H os covering systems . J. Eur. Math. Soc. (JEMS) (2024), 75--109.

[Er80] Erd\H{o}s, Paul, A survey of problems in combinatorial number theory . Ann. Discrete Math. (1980), 89-115.

[Ho15] Hough, Bob, Solution of the minimum modulus problem for covering systems . Ann. of Math. (2) (2015), 361-382.

\noindent\textbf{Related OEIS sequences:} possible


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1188}}
\end{document}
