\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1112}
\date{Last updated: 2025-12-28}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} additive combinatorics

\medskip

\noindent\textbf{Problem Statement:}

Let $1\leq d_1&#60;d_2$ and $k\geq 3$. Does there exist an integer $r$ such that if $B=\{b_1&#60;\cdots\}$ is a lacunary sequence of positive integers with $b_{i+1}\geq rb_i$ then there exists a sequence of positive integers $A=\{a_1&#60;\cdots\}$ such that\[d_1\leq a_{i+1}-a_i\leq d_2\]for all $i\geq 1$ and $(kA)\cap B=\emptyset$, where $kA$ is the $k$-fold sumset?




Erd\H{o}s and Graham \cite{ErGr80} noted that if $B=\{b_1&#60;b_2&#60;\cdots\}$ with $b_1\geq 5$ and $b_{i+1}\geq 2b_i$ then there is a set $A=\{a_1&#60;a_2&#60;\cdots\}$ with $2\leq a_{k+1}-a_k\leq 3$ for all $k$ such that $(A+A)\cap B=\emptyset$. Bollob\'{a}s, Hegyv\'{a}ri, and Jin \cite{BHJ97} showed that such an $A$ must exist if $b_{i+1}\geq 2b_i-O(1)$, and that this is best possible. Erd\H{o}s and Graham go on to say 'whether such behavior can hold for $A+A+A$ (or more summands) is not known'. The above question is my best interpretation of what they intended. Bollob\'{a}s, Hegyv\'{a}ri, and Jin \cite{BHJ97} provide a negative answer in that, for any sequence of integers $1\leq r_1&#60;r_2&#60;\cdots$, there is a $B$ as above with $b_{i+1}\geq r_ib_i$ such that $(A+A+A)\cap B\neq\emptyset$ for any $A$ with $2\leq a_{i+1}-a_i\leq 3$. They define, more generally, $r_k(d_1,d_2)$ as the smallest $r$ (if it exists) such that if $b_{i+1}\geq rb_i$ then there exists $A$ with $d_1\leq a_{i+1}-a_i\leq d_2$ such that $(kA)\cap B=\emptyset$, where $kA$ is the $k$-fold sumset. It follows from the above results that $r_2(2,3)=2$ and that $r_3(2,3)$ does not exist. Chen \cite{Ch00} proved that $r_2(a,b)\leq 2$ for any integers $a&#60;b$ with $b\neq 2a$, and that $r_2(a,2a)\geq 2$ for all integers $a$. The more general question of existence of $r_k(a,b)$ for $k\geq 3$ remains open. Some further technical non-existence results are given by Tang and Yang \cite{TaYa21}. [Note the stated problem is a generous interpretation of a very ambiguous remark in \cite{ErGr80}, so it might be more appropriate to call this a problem 'inspired by Erd\H{o}s and Graham'.]




References

[BHJ97] Bollob\'{a}s, B\'{e}la and Hegyv\'{a}ri, Norbert and Jin, Guoping, On a problem of {E}rd\H{o}s and {G}raham . Discrete Math. (1997), 253--257.

[Ch00] Chen, Yong-Gao, On sums and intersects of sequences . Discrete Math. (2000), 351--354.

[ErGr80] Erd\H{o}s, P. and Graham, R., Old and new problems and results in combinatorial number theory . Monographies de L'Enseignement Mathematique (1980).

[TaYa21] Tang, Min and Yang, Quan-Hui, On a problem of {E}rd\H{o}s and {G}raham . Publ. Math. Debrecen (2021), 485--493.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1112}}
\end{document}
