\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#208}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory

\medskip

\noindent\textbf{Problem Statement:}

Let $s_1&lt;s_2&lt;\cdots$ be the sequence of squarefree numbers. Is it true that, for any $\epsilon&gt;0$ and large $n$,\[s_{n+1}-s_n \ll_\epsilon s_n^{\epsilon}?\]Is it true that\[s_{n+1}-s_n \leq (1+o(1))\frac{\pi^2}{6}\frac{\log s_n}{\log\log s_n}?\]




Erd\H{o}s \cite{Er51} showed that there are infinitely many $n$ such that\[s_{n+1}-s_n &gt; (1+o(1))\frac{\pi^2}{6}\frac{\log s_n}{\log\log s_n},\]so this bound would be the best possible. In \cite{Er79} Erd\H{o}s says perhaps $s_{n+1}-s_n \ll \log s_n$, but he is 'very doubtful'. Filaseta and Trifonov \cite{FiTr92} proved an upper bound of $s_n^{1/5+o(1)}$. Pandey \cite{Pa24} has improved this exponent to $1/5-c$ for some constant $c&gt;0$. Granville \cite{Gr98} showed that $s_{n+1}-s_n\ll_\epsilon s_n^\epsilon$ for all $\epsilon&gt;0$ follows from the ABC conjecture . See also [489] and [145] . A more general form of this problem is given in [1101] .




References

[Er51] Erd\H{o}s, P., Some problems and results in elementary number theory . Publ. Math. Debrecen (1951), 103-109.

[Er79] Erd\H{o}s, Paul, Some unconventional problems in number theory . Math. Mag. (1979), 67-70.

[FiTr92] Filaseta, M. and Trifonov, O., On gaps between squarefree numbers II . J. London Math. Soc. (1992), 215-221.

[Gr98] Granville, Andrew, {$ABC$} allows us to count squarefrees . Internat. Math. Res. Notices (1998), 991--1009.

[Pa24] Pandey, M., Squarefree numbers in short intervals . arXiv:2401.13981 (2024).

\noindent\textbf{Related OEIS sequences:} A005117, A076259


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/208}}
\end{document}
