\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#883}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, graph theory

\medskip

\noindent\textbf{Problem Statement:}

For $A\subseteq \{1,\ldots,n\}$ let $G(A)$ be the graph with vertex set $A$, where two integers are joined by an edge if they are coprime. Is it true that if\[\lvert A\rvert &#62;\lfloor\tfrac{n}{2}\rfloor+\lfloor\tfrac{n}{3}\rfloor-\lfloor\tfrac{n}{6}\rfloor\]then $G(A)$ contains all odd cycles of length $\leq \frac{n}{3}+1$? Is it true that, for every $\ell\geq 1$, if $n$ is sufficiently large and\[\lvert A\rvert &#62;\lfloor\tfrac{n}{2}\rfloor+\lfloor\tfrac{n}{3}\rfloor-\lfloor\tfrac{n}{6}\rfloor\]then $G(A)$ must contain a complete $(1,\ell,\ell)$ triparite graph on $2\ell+1$ vertices?




A problem of Erd\H{o}s and S\'{a}rk\H{o}zy \cite{ErSa97}, who prove that if\[\lvert A\rvert &#62;\lfloor\tfrac{n}{2}\rfloor+\lfloor\tfrac{n}{3}\rfloor-\lfloor\tfrac{n}{6}\rfloor\]then $G(A)$ contains all odd cycles of length $\leq cn$ for some constant $c&#62;0$. This threshold is the best possible, since one could take $A$ to be the set of $m\leq n$ which are divisible by either $2$ or $3$, in which case $G(A)$ contains no triangles. The second question was solved by S\'{a}rk\"{o}zy \cite{Sa99} who proved that, for large $n$, if $\lvert A\rvert$ exceeds the given threshold then $G(A)$ contains a complete $(1,\ell,\ell)$ triparite graph with\[\ell \gg \frac{\log n}{\log\log n}.\]




References

[ErSa97] Erd\H{o}s, Paul and Sarkozy, Gabor N., On cycles in the coprime graph of integers . Electron. J. Combin. (1997), Research Paper 8, approx. 11.

[Sa99] S\'{a}rk\"ozy, G\'{a}bor N., Complete tripartite subgraphs in the coprime graph of integers . Discrete Math. (1999), 227--238.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/883}}
\end{document}
