\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#1057}
\date{Last updated: 2025-09-28}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory

\medskip
\noindent\textbf{Note:} Carmichael numbers

\medskip

\noindent\textbf{Problem Statement:}

Let $C(x)$ count the number of Carmichael numbers in the interval $[1,x]$. Is it true that $C(x)=x^{1-o(1)}$?




Erd\H{o}s \cite{Er56c} proved\[C(x) &lt; x \exp\left(-c \frac{\log x\log\log\log x}{\log\log x}\right)\]for some constant $c&gt;0$. Pomerance \cite{Po89} gave a heuristic suggesting that this is the true order of growth, and in fact\[C(x)= x \exp\left(-(1+o(1))\frac{\log x\log\log\log x}{\log\log x}\right).\]Alford, Granville, and Pomerance \cite{AGP94} proved that $C(x)\to \infty$, and in fact $C(x)&#62;x^{2/7}$ for large $x$. The lower bound\[C(x)&#62; x^{0.33336704}\]was proved by Harman \cite{Ha08}. This exponent was improved to $0.3389$ by Lichtman \cite{Li22}. Korselt observed that $n$ being a Carmichael number is equivalent to $n$ being squarefree and $p-1\mid n-1$ for all primes $p\mid n$. This is discussed in problem A13 of Guy's collection \cite{Gu04}.




References

[AGP94] Alford, W. R. and Granville, Andrew and Pomerance, Carl, There are infinitely many {C}armichael numbers . Ann. of Math. (2) (1994), 703--722.

[Er56c] Erd\H{o}s, P., On pseudoprimes and {C}armichael numbers . Publ. Math. Debrecen (1956), 201--206.

[Gu04] Guy, Richard K., Unsolved problems in number theory . (2004), xviii+437.

[Ha08] Harman, Glyn, Watt's mean value theorem and {C}armichael numbers . Int. J. Number Theory (2008), 241--248.

[Li22] J. D. Lichtman, Primes in arithmetic progressions to large moduli and shifted primes without large prime factors . arXiv:2211.09641 (2022).

[Po89] Pomerance, Carl, Two methods in elementary analytic number theory . (1989), 135--161.

\noindent\textbf{Related OEIS sequences:} A006931


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/1057}}
\end{document}
