\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#501}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} combinatorics, set theory

\medskip

\noindent\textbf{Problem Statement:}

For every $x\in\mathbb{R}$ let $A_x\subset \mathbb{R}$ be a bounded set with outer measure $&#60;1$. Must there exist an infinite independent set, that is, some infinite $X\subseteq \mathbb{R}$ such that $x\not\in A_y$ for all $x\neq y\in X$? If the sets $A_x$ are closed and have measure $&#60;1$, then must there exist an independent set of size $3$?




Erd\H{o}s and Hajnal \cite{ErHa60} proved the existence of arbitrarily large finite independent sets (under the assumptions in the first problem). Gladysz \cite{Gl62} proved the existence of an independent set of size $2$ under the assumptions of the second question. Hechler \cite{He72} has shown the answer to the first question is no, assuming the continuum hypothesis. Newelski, Pawlikowski, and Seredy\'{n}ski \cite{NPS87} proved that if all the $A_x$ are closed with measure $&#60;1$ then there is an infinite independent set, giving a strong affirmative answer to the second question.




References

[ErHa60] Erd\H{o}s, P. and Hajnal, A., Some remarks on set theory. VIII . Michigan Math. J. (1960), 187-191.

[Gl62] G\l adysz, S., Bemerkungen \"uber die {U}nabh\"{a}ngigkeit der {P}unkte in {B}ezug auf mengenwertige {F}unktionen . Acta Math. Acad. Sci. Hungar. (1962), 199--201.

[He72] Hechler, S. H., On two problems in combinatorial set theory . Bull. Acad. Polon. Sci. S\'{e}r. Sci. Math. Astronom. Phys. (1972), 429-431.

[NPS87] Newelski, Ludomir and Pawlikowski, Janusz and Seredy\'nski, Witold, Infinite free set for small measure set mappings . Proc. Amer. Math. Soc. (1987), 335--339.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/501}}
\end{document}
