\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#267}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} irrationality

\medskip

\noindent\textbf{Problem Statement:}

Let $F_1=F_2=1$ and $F_{n+1}=F_n+F_{n-1}$ be the Fibonacci sequence. Let $n_1&lt;n_2&lt;\cdots $ be an infinite sequence with $n_{k+1}/n_k \geq c&gt;1$. Must\[\sum_k\frac{1}{F_{n_k}}\]be irrational?




It may be sufficient to have $n_k/k\to \infty$. Good \cite{Go74} and Bicknell and Hoggatt \cite{BiHo76} have shown that $\sum \frac{1}{F_{2^n}}$ is irrational - in fact,\[\sum \frac{1}{F_{2^n}}=\frac{7-\sqrt{5}}{2}.\]Badea \cite{Ba87} proved that $\sum \frac{1}{F_{2^n+1}}$ is irrational. The sum $\sum \frac{1}{F_n}$ itself was proved to be irrational by Andr\'{e}-Jeannin \cite{An89}. The main problem has been proved for $c\geq 2$ by Badea \cite{Ba93}. It remains open for $1&#60;c&#60;2$.




References

[An89] Andr\'{e}-Jeannin, Richard, Irrationalit\'{e}{} de la somme des inverses de certaines suites r\'{e}currentes . C. R. Acad. Sci. Paris S\'{e}r. I Math. (1989), 539--541.

[Ba87] Badea, C., The irrationality of certain infinite series . Glasgow Math. J. (1987), 221--228.

[Ba93] Badea, C., A theorem on irrationality of infinite series and applications . Acta Arith. (1993), 313--323.

[BiHo76] Hoggatt, Jr., V. E. and Bicknell, Marjorie, A reciprocal series of Fibonacci numbers with subscripts $2^nk$ . Fibonacci Quart. (1976), 453-455.

[Go74] Good, I. J., A reciprocal series of Fibonacci numbers . Fibonacci Quart. (1974), 346.


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/267}}
\end{document}
