\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#513}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} analysis

\medskip

\noindent\textbf{Problem Statement:}

Let $f=\sum_{n=0}^\infty a_nz^n$ be a transcendental entire function. What is the greatest possible value of\[\liminf_{r\to \infty} \frac{\max_n\lvert a_nr^n\rvert}{\max_{\lvert z\rvert=r}\lvert f(z)\rvert}?\]




Let $B$ be the supremum of all such values. It is trivial that $B\in [1/2,1]$. K\"{o}v\'{a}ri (unpublished) observed that it must be $&gt;1/2$. Gray and Shah \cite{GrSh63} give an argument of Clunie which shows $B\leq 2/\pi$. Clunie and Hayman \cite{ClHa64} improved both bounds to\[4/7&lt; B \leq 2/\pi-c\]for some absolute constant $c&gt;0$. He and Tang \cite{HeTa26} have improved the lower bound to\[B&gt; 0.5850724\](note that $4/7\approx 0.57143$). This was further improved slightly to $0.5850788$ by GPT (as prompted by Sothanaphan). Note that $2/\pi \approx 0.63662$. See also [227] .




References

[ClHa64] Clunie, J. and Hayman, W. K., The maximum term of a power series . J. Analyse Math. (1964), 143-186.

[GrSh63] Gray, Alfred and Shah, S. M., A note on entire functions and a conjecture of Erd\H{o}s . Bull. Amer. Math. Soc. (1963), 573-577.

[HeTa26] Y. He and Q. Tang, Generalizing the Clunie--Hayman construction in an Erd\H{o}s maximum-term problem . arXiv:2602.12217 (2026).


\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/513}}
\end{document}
