\documentclass{article}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}

\title{Erd\H{o}s Problem \#870}
\date{Last updated: 2025-08-31}
\author{Source: erdosproblems.com}

\begin{document}
\maketitle

\noindent\textbf{Status:} OPEN \quad \textbf{No prize}

\noindent\textbf{Tags:} number theory, additive basis

\medskip
\noindent\textbf{Problem Statement:}

Let $k\geq 3$ and $A$ be an additive basis of order $k$. Does there exist a constant $c=c(k)>0$ such that if $r(n)\geq c\log n$ for all large $n$ then $A$ must contain a minimal basis of order $k$? (Here $r(n)$ counts the number of representations of $n$ as the sum of at most $k$ elements from $A$.)

\bigskip
\noindent\textbf{Remarks:}

A question of Erd\H{o}s and Nathanson \cite{ErNa79}, who proved that this is true for $k=2$ if $1_A\ast 1_A(n) > (\log \frac{4}{3})^{-1}\log n$ for all large $n$.

Härtter \cite{Ha56} and Nathanson \cite{Na74} proved that there exist additive bases which do not contain any minimal additive bases. 

See also [868].

\bigskip
\noindent\textbf{References:}

[ErNa79] Erd\H{o}s, Paul and Nathanson, Melvyn B., Systems of distinct representatives and minimal bases in
additive number theory. (1979), 89--107.
 
 [Ha56] H\"{a}rtter, Erich, Ein Beitrag zur Theorie der {M}inimalbasen. J. Reine Angew. Math. (1956), 170--204.
 
 [Na74] Nathanson, Melvyn B., Minimal bases and maximal nonbases in additive number theory. J. Number Theory (1974), 324--333.

\bigskip
\noindent\small{Source: \url{https://www.erdosproblems.com/870}}
\end{document}
